---
title: "A raid in BiciMad"
author: "@vsuarezlledo"
date: 2018-04-23
categories: ["R"]
tags: ["r", "data", "public"]
math: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

# BiciMad `r emo::ji("bike")`

The [Open Data Portal of Madrid City Council](http://datos.madrid.es/portal/site/egob) is great initiative also is a good place to have enough data to try. Insipired by [this post](http://toddwschneider.com/posts/a-tale-of-twenty-two-million-citi-bikes-analyzing-the-nyc-bike-share-system/) I will dig into extracting the maximum information possible. For this porpouse I will use the [available data of BiciMad](http://datos.madrid.es/egob/catalogo/300110-0-accidentes-bicicleta.csv). All code is available on [GitHub](https://github.com/VictorSuarezL/vsslledo).

Packages needed:
```{r message=FALSE, warning=FALSE, echo= F}
library(httr)
library(readr)
library(tidyverse)
library(emo)
library(scales)
```

Let's get the data. First of all set the `url` hosting data. Latter `read_delim` will help us with this task, set semicolon as delimiter then, just in case, `trim_ws = T` will trim whitespace.

```{r}
url <- "http://datos.madrid.es/egob/catalogo/300110-0-accidentes-bicicleta.csv"

csv_data <- read_delim(url, delim = ";", trim_ws = T, locale = locale(encoding = "latin1"))

# El sistema operativo es UTF-8 y el encode original del archivo es latin1
# csv_data <- read.csv(url, sep = ";", header = T, skip = 1, strip.white = T, stringsAsFactors = F, encoding = "UTF-8", fileEncoding = "latin1")
```

##**Where are there more accidents with bicycles?**

To answer this question we are going to use a couple of resources inside the `ggmap` package. Through this package we can access different map sources. By adjusting the `get_map` function we will have a map of the Community of Madrid.

```{r cache.lazy=TRUE, include=FALSE}
library(ggmap)

m <- get_map("Comunidad de Madrid, España", zoom = 13, maptype = "terrain-lines")
```

Once we have the map we must first search the places, in general, street names. We also add the number of this one and in second place we add `" Madrid "` to ensure that the location is within the Community of Madrid. To obtain the location, we must simply use the `geocode` function which returns the coordinates of the address.

```{r eval=FALSE}
madrid_sites <- paste(csv_data$Lugar, csv_data$Numero, "Madrid", sep = " ")

gcode <- function(x) {
  Sys.sleep(5)
  geocode(x)
}

xx <- map_df(madrid_sites, gcode)

#Just erasing Madrid from madrid_sites
xx$location <- sub("Madrid", "", madrid_sites)
#Just cleaning
xx$location <- tm::stripWhitespace(xx$location)

write_csv(xx, path = "./data/geocodes.csv")
```
```{r message=FALSE, warning=FALSE, include=FALSE}
xx <- read_csv("/Users/Tito/Documents/GitHub/vsslledo/data/geocodes.csv")
```

The final step is to gather everything we have obtained so far. On the one hand the map and on the other hand the `geocode`:

```{r fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
## Map with locations:
ggmap(m) +
  geom_point(data = xx, aes(fill = location), size = 2, col = "#ff7058") +
  coord_map() + 
  guides(fill=FALSE) + labs(x = "", y = "") +
  theme_void() + ggtitle("")
```